---
title: "Titanic survival prediction"
date: "11 april 2018"
author: "Eliza Norden, Alex Moellers, Sarah Catherine James"
output:
  pdf_document: default
  html_document: 
    keep_md: true
---

# Read data and typecast relevant variables
``` {r}
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(ggplot2)
library(randomForest)

train = read.csv("train.csv")
test = read.csv("test.csv")

# combine train and test data
test$Survived = rep("None",418)
alldata = rbind(train, test)
attach(alldata)
train$Pclass <- as.factor(train$Pclass)
train$Survived <- as.factor(train$Survived)
alldata$Survived <- as.factor(alldata$Survived)
alldata$Pclass <- as.factor(alldata$Pclass)

```

# Visualize data
```{r}
table(alldata$Survived) # Much more people died than survived
table(alldata$Pclass)  # Second and First class have a similar amount of people, third has many. 

ggplot(data = train, aes(x = Pclass, fill = Survived )) + geom_histogram(stat = "count")
ggplot(data = train, aes(x = Sex, fill = Survived )) + geom_histogram(stat = "count")

```
In total, survival rate 62%. Most passengers onboard stayed in third class.
We can see that these had a lower probability of surviving than people from the other classes. We also see a strong survival bias among females. 

# Feature extraction #1 - titles
```{r}

# Let us deal with the name feature
length(unique(as.character(alldata$Name)))  # Are there some names occurring two times?
double.names <- as.character(alldata[which(duplicated(as.character(alldata$Name))),"Name"])
double.names.entries <- alldata[which(as.character(alldata$Name) == double.names[1]|as.character(alldata$Name) == double.names[2]),]

# Some people seem to have the same name, but they are not the same person. 
# Now let us deal with the titles.

Title = sapply(Name, FUN = function(x) {strsplit(as.character(x), split='[,.]')[[1]][2]})
Title = substring(Title, 2)
alldata$Title = as.character(Title)

# That is a lot of titles, lets group them together.

alldata$Title[alldata$Title %in% c("Don", "Jonkheer")] = "Sir"
alldata$Title[alldata$Title %in% c("the Countess", "Lady", "Dona")] = "Lady"
alldata$Title[alldata$Title %in% c("Col", "Capt","Major")] = "Military"

# Let us not overfit, the seem to be the same as their sex and passenger class

alldata$Title[alldata$Title=="Mlle" | alldata$Title=="Ms"] = "Miss" # We are assuming here that all Ms are unmarried.
alldata$Title[alldata$Title=="Mme"] = "Mrs"

ggplot(data = alldata[1:891,], aes(x= Title, fill = Survived)) + geom_histogram(stat = "count")

# Let us not overfit, they seem to be the same as their sex and passenger class

ind_Tit_F = which(alldata$Title =="Lady")
alldata$Title[ind_Tit_F] = "Mrs"

ind_Tit_Male = which(alldata$Title =="Rev"|alldata$Title =="Military"|alldata$Title =="Dr"|alldata$Title =="Sir")
alldata$Title[ind_Tit_Male] = "Mr"

alldata$Title = as.factor(alldata$Title)

# Do these titles hold any predicitve power?
ggplot(data = alldata[1:891,], aes(x= Title, fill = Survived)) + geom_histogram(stat = "count")

# Now let us make the plot again and also include the Pclass.
ggplot(data = alldata[1:891,], aes(x= Title, fill = Survived)) + geom_histogram(stat = "count", binwidth = 0.6) + facet_wrap("Pclass")

```


# Feature extraction #2 - family size
``` {r}

# Let us turn to the sibspa variable. There is seven different levels of this variable. We will convert it into a factor and take a look at its predictive power.

ggplot(alldata[1:891,], aes(x = as.factor(SibSp), fill = Survived)) + geom_histogram(stat= "count", binwidth = 0.6) + facet_wrap(~Pclass+Title)

# We note that travelling with fewer people increases the survival rate in some cases. Especially for Master and Miss. To be able to take this into account, we create a family size variable. 

temp.sibsp <- c(train$SibSp, test$SibSp)
temp.parch <- c(train$Parch, test$Parch)
alldata$FamSize <- as.factor(temp.sibsp+temp.parch+1)

ggplot(alldata[1:891,], aes(x = FamSize, fill = Survived)) + geom_histogram(stat= "count", binwidth = 0.6) + facet_wrap(~Pclass+Title)

```
Passengers holding the title Mr seem to be likely to die, regardless of passenger class, numbers of siblings/spouses or total family size. However, first class Mr's traveling with a small family seems to have improved survival rate over solo travelers and big families.

# Feature extraction #3 - boat decks and individual passenger fares
``` {r}
# create decks
alldata$Deck = substring(Cabin, 1, 1)
alldata$Deck[alldata$Deck==""] = "U" #unknown
alldata$Deck = as.factor(alldata$Deck)

ggplot(data = alldata,aes(x=as.factor(alldata$Survived), fill=alldata$Title)) + geom_bar(stat="count") 

#alldata$FamSize = as.factor(alldata$SibSp + alldata$Parch)
alldata$FamSize = alldata$SibSp + alldata$Parch

# Fare previous for the whole ticket. Divide it by number of passengers per ticket to get fare per passenger. 
# Travel size defined by number of passengers sharing same ticket number. Similar to FamSize but takes friend groups into account
alldata$FareAdj = rep("", dim(alldata)[1])
alldata$TravelSize = (rep(1, dim(alldata)[1]))
for(ticnr in levels(Ticket)) {
  alldata$FareAdj[alldata$Ticket==ticnr] = signif(alldata$Fare[(alldata$Ticket==ticnr)]/sum(alldata$Ticket==ticnr),4)
  alldata$TravelSize[(alldata$Ticket==ticnr)] =(sum(alldata$Ticket==ticnr))
}

alldata$FareAdj = as.numeric(alldata$FareAdj)


``` 
Fare adjusted for each passenger should be a better predictor than total ticket fare, since the fare data is more consistent. Instead, the size of the traveling company has been captured in its own variable.


# Data imputation
```{r}

summary(alldata)
# One NA adjusted fare. Estimate it using a regression tree.
FareFit = rpart(FareAdj ~ Sex + Pclass + Title + Deck + Embarked + TravelSize + FamSize, data=alldata[!is.na(alldata$FareAdj),], method="anova")
alldata$FareAdj[is.na(alldata$FareAdj)] = predict(FareFit, newdata=alldata[is.na(alldata$FareAdj),])
# Fill in the corresponding ticket fare for completeness.
alldata$Fare[is.na(alldata$Fare)] = alldata$FareAdj[is.na(alldata$Fare)]

# 2 missing embarkation ports. Use FareAdj as predictor.
EmbarkedFit = rpart(Embarked ~ FareAdj, data=alldata[alldata$Embarked!="",], method="class", maxdepth=3)
alldata$Embarked[which(alldata$Embarked=="")] = predict(EmbarkedFit, newdata=alldata[alldata$Embarked=="",], type="class")

# 263 NA ages. Regression by random forest.
AgeFit = randomForest(Age ~ FareAdj + Title + FamSize, data=alldata[!is.na(alldata$Age),], importance=TRUE, ntrees=500)
alldata$Age[is.na(alldata$Age)] = round(predict(AgeFit, newdata=alldata[is.na(alldata$Age),]))

```

# Model fit and prediction
``` {r} 
# Cross validation
library(ggplot2)
library(caret)
library(doSNOW)
library(e1071)

# Final choice of predictors
train.set.1 = alldata[1:891, c("Pclass", "Title","FareAdj","TravelSize")]
train.set.1$Title <- as.factor(train.set.1$Title)
rf.label <- as.factor(train$Survived)

set.seed(5000)
# Split data into folds for cross validation
cross.10 <- createMultiFolds(rf.label, k=3, times = 10)
ctrl.1 <- trainControl(method = "repeatedcv", number = 10, repeats = 10, index = cross.10)

set.seed(2100)
# Cross validation
cv.result.1 <- train(x = train.set.1, y = rf.label, method = "rf", tuneLength = 3, ntree = 1000, trControl = ctrl.1)
#this will take some time
cv.result.1

# Best model and number of sampled variables retrieved from cross validation
final.model = cv.result.1$finalModel
mtry = cv.result.1$bestTune

test.set = alldata[892:1309,  c("Pclass", "Title","FareAdj","TravelSize")]
pred = predict(final.model, test.set, mtry=mtry)

df = data.frame(PassengerId = alldata$PassengerId[892:1309], Survived = pred)
write.csv(df, "submit_cv.csv", row.names=FALSE)

```
The best random forest model uses 2 randomly sampled variables at each split. Predicted test accuracy is 0.833. The actual test accuracy, retrieved by prediction submission to the Kaggle competition, is 0.799.


